Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-22 ICPP-19/communications.tex
Proofread

	In this class of strategies, we precompute the order of tasks based on some criterion and then follow this ordering as much as possible. But when the communication resource is idle because the memory requirement of the next task is too high, then we select a task  with a dynamic strategy. After a task is dynamically selected, we update the remaining order without this task. This class of strategies take advantage of static information in the form of precomputed order and dynamic correction to minimize the idle time due to memory constraint.

	In all strategies of this class, the initial order is $OMIM$ order,  obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. We define the following heuristics based on how we select a task from the set of tasks which fit in memory and induce minimum idle time on the computation resource. If no task fits in memory then we leave the resource idle and forward to the next event point.
	%%	\todo[inline]{LED: What happens if no task fit in memory?}

@@

	\subsection{Solving Linear Program Iteratively}
	\label{subsec:linearprogrammingformulation}
	We use a mixed integer linear program to obtain the order of communications and computations. Recall that $COMP_i$ and $COMM_i$ represent computation and communication times of task $i$, and the memory capacity of the target system is $C$. In the linear program formulation, $s_i$ and $e_i$ (resp. $s'_i$ and $e'_i$) represent the start and end times of communication (resp. computation) for task $i$, and $MC(i)$ is the memory capacity requirement of task $i$. The formulation also contains for each pair of tasks $i$ and $j$ \textit{i)} a boolean variable $a_{ij}$ to denote the order of $i$ and $j$ on the communication resource \textit{ii)} a boolean variables $b_{ij}$ to denote the order of $i$ and $j$ on the computation resource, and \textit{iii)} a boolean variables $c_{ij}$ to denote the order of $s_i$ and $e'_j$.

@@

		In this paper, we consider the problem of deciding the order of data transfers between two memory nodes such that overlap of communications and computations is maximized. With Exascale computing, applications face bottlenecks due to communications. Hence, it is extremely important to achieve the maximum  communication-computation overlap in order to exploit the full potential of the system. We show that determining the order of data transfers is a NP complete problem. We propose several data transfer heuristics and evaluate them on two molecular chemistry kernels, HF and CCSD. Our results show that some of our heuristics achieve significant overlap and their makespans are very close to the lower bound. We plan to evaluate our strategies on different applications coming from multiple domains. We also plan to study the behavior of our strategies in the context of overlapping CPU-GPU communications with computations. A runtime system aiming to expose different heuristics to maximize the communication-computation overlap at the developer level and automatically select the best one is currently underway.

>>>>

	In this class of strategies, we precompute the order of tasks based on some criterion and then follow this ordering as much as possible. But when the communication resource is idle because the memory requirement of the next task is too high, then we select a task  with a dynamic strategy. After a task is dynamically selected, we update the remaining order without this task. This class of strategies takes advantage of static information in the form of precomputed order and dynamic correction to minimize the idle time due to memory constraint.

	In all strategies of this class, the initial order is $OMIM$ order obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. We define the following heuristics based on how we select a task from the set of tasks which fit in memory and induce minimum idle time on the computation resource. If no task fits in memory then we leave the resource idle and forward to the next event point.
	%%	\todo[inline]{LED: What happens if no task fit in memory?}

@@

	\subsection{Solving Linear Program Iteratively}
	\label{subsec:linearprogrammingformulation}
	We use a mixed integer linear program to obtain the order of communications and computations. Recall that $COMP_i$ and $COMM_i$ represent computation and communication times of task $i$, and the memory capacity of the target system is $C$. In the linear program formulation, $s_i$ and $e_i$ (resp. $s'_i$ and $e'_i$) represent the start and end times of communication (resp. computation) for task $i$, and $MC(i)$ is the memory capacity requirement of task $i$. The formulation also contains for each pair of tasks $i$ and $j$ \textit{i)} a boolean variable $a_{ij}$ to denote the order of $i$ and $j$ on the communication resource \textit{ii)} a boolean variable $b_{ij}$ to denote the order of $i$ and $j$ on the computation resource, and \textit{iii)} a boolean variable $c_{ij}$ to denote the order of $s_i$ and $e'_j$.

@@

		In this paper, we consider the problem of deciding the order of data transfers between two memory nodes such that overlap of communications and computations is maximized. With Exascale computing, applications face bottlenecks due to communications. Hence, it is extremely important to achieve the maximum  communication-computation overlap in order to exploit the full potential of the system. We show that the problem of determining the optimal order of data transfers is NP complete. We propose several data transfer heuristics and evaluate them on two molecular chemistry kernels, HF and CCSD. Our results show that some of our heuristics achieve significant overlap and their makespans are very close to the lower bound. We plan to evaluate our strategies on different applications coming from multiple domains. We also plan to study the behavior of our strategies in the context of overlapping CPU-GPU communications with computations. A runtime system aiming to expose different heuristics to maximize the communication-computation overlap at the developer level and automatically select the best one is currently underway.
